---
title: "Test constructions and test scores"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(psych)
library(gradethis)
library(ggplot2)
library(CTT)
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE)

```

## 1. Exam Data (MC-items)
A data frame with 100 observations on the following 20 items. 

```{r, source, exercise=TRUE, echo=FALSE}
data(CTTdata)
attach(CTTdata) 
head(CTTdata)
data(CTTkey)

data <- score(CTTdata,CTTkey, output.scored=TRUE)
response <- as.data.frame(data$scored)
head(response) # this saves the item score (0 or 1)
```

## 2. Item Difficulty 
- The proportion of those responding correctly

```{r,  data, exercise=TRUE, echo=FALSE, exercise.setup="source"}

# Let's calculate the difficulty of item 2

p = sum(response$i2)/100
print(p)

```

```{r,  data2, exercise=TRUE, echo=FALSE, exercise.completion=TRUE, exercise.setup="source"}

# Let's calculate the difficulty of item 10



```

## 3. Inter-item Reliability 
- inter-item correlations from joint frequency distribution tables

```{r,  data3, exercise=TRUE, echo=FALSE, exercise.setup="source"}
# Let's compute the inter-item correlation score between item 7 and 9 
# First, we will have to create a joint frequency distribution table 

table(response$i7, response$i9)

```

```{r, data4, exercise=TRUE, echo=FALSE, exercise.lines=15, exercise.setup="source"}

# Second, we will have to identify the following components (p16, Topic 2)

pjk = 
  
pj = 
  
pk = 
  
qj = 
  
qk = 

# The final outcome will be: 
rho = (pjk-(pj*pk))/(sqrt(pj*qj*pk*qk))  
```


## 4. Total Test scores 
Total test scores are frequently used in low stakes assessment

```{r, data5, exercise=TRUE, echo=FALSE, exercise.lines=10, exercise.setup="source"}
# total score can be computed using rowSums()

rowSums(response)[13]

###### YOUR CODE HERE #########


################################
```

```{r letter-b, echo=FALSE}

question("What is the total test score of the examinee 11?",
  answer("10", correct = TRUE),
  answer("11"),
  answer("12"), 
  answer("13")
)
```

Let's compute the average test score and compare it with the total item difficulty 
```{r, data6, exercise=TRUE, echo=FALSE, exercise.lines=16, exercise.setup="source"}
# 1. Average Test Score 
###### YOUR CODE HERE #########
ave_test_score = 
################################

# 2. Sum of all item difficulty
###### YOUR CODE HERE #########
sum_dif = 
################################

# 3. Check whether ave_test_score == sum_dif 
###### YOUR CODE HERE #########

################################

```
