---
title: "CTT Item Analysis"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(psych)
library(gradethis)
library(truncnorm)
library(ggplot2)
library(CTT)
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE)

```

## 1. Exam Data (MC-items)
A data frame with 100 observations on the following 20 items. 

```{r data2, exercise=TRUE, echo=FALSE} 

data(CTTdata)
attach(CTTdata) 
head(CTTdata)

print('Answer key is:')
data(CTTkey)
print(CTTkey) 
```

### Change the dataframe to dichotomous 
```{r, score_data, echo=TRUE, execise.setup="data2"}

data <- score(CTTdata,CTTkey, output.scored=TRUE)
response <- as.data.frame(data$scored)
scores <- as.data.frame(data$score)

head(scores) # this saves each student's total score 
head(response) # this saves the item score (0 or 1)
```

### Test Score 
Let's use `describe()` to look at the descriptive statistics of the `scores` data

```{r, test, echo=FALSE, exercise=TRUE, execise.setup="score_data"}
describe(scores)

x <- scores$'data$score'
h <- hist(x, las=1, breaks="FD", xlab= "Red vertical line shows the mean.", 
          main="", col = "cyan")
rug(x)
abline(v = mean(x), col = "red", lwd = 2)

```

## 2. Item Difficulty 

Now we have a new data frame called `response` which stores the dichotomous item scores
```{r, response_data, echo=TRUE, exercise.setup="score_data"} 
head(response)
```

### Item Difficulty (P value)

In CTT, item difficulty, also known as item p-value, is often quantified in terms of the mean item score or the percentage of students who answer the item correctly. 

Let's calculate the P-value using `colMeans()`

```{r, answer1, echo=TRUE, exercise=TRUE, exercise.setup="score_data"} 
# Let's compute the p-value first (and call this `pj`)
pj = # compute the p values 

```

```{r, answer1-solution}
# Let's compute the p-value first 
pj = colMeans(response) 

assign("pj", pj, envir=globalenv()) 
```

```{r, answer1-check}
grade_this_code()
```

- Q1: Which of the following items is the most difficult?
- Q2: Which of the following items is the least difficult? 

## 3. Item Discrimination 
Item discrimination can be quantified in terms of the item-total correlation. 

- Q1: What types of discrimination index should we adopt? 

- Point-biserial correlation:
$$ \rho_{pbis} = [\frac{\mu_{+}-\mu_{X}}{\sigma_{X}}](\sqrt\frac{p}{q})$$ 
- Spearman’s rho correlation:
$$ d_{j} = \frac{\sum_{i=1}^{N}(X_{ij}-p_{j})(X_{C_{i}} - \mu_{X_{c}})}{\sqrt{\sum_{i=1}^{N}(X_{ij}-p_{j})^2\sum_{i=1}^{N}{(X_{C{i}}-\mu_{X_{c}})^2}}} $$ 

### Item Analysis ``itemAnalysis()``


```{r, IA1, echo=TRUE, exercise=TRUE, exercise.setup="score_data"} 
# Let's compute item discrimination for item 10 (i10):
corrected_data = subset(response, response$i10==1)

m1 = mean(rowSums(subset(corrected_data, select = -c(i10) )))
mt = mean(rowSums(subset(response, select = -c(i10) )))
sd_t = sd(rowSums(subset(response, select = -c(i10) )))
p = sum(response$i10)/100
q = 1- p

pbis = ((m1-mt)/sd_t)*(sqrt(p/q))
```

```{r, IA, echo=TRUE, exercise=TRUE, exercise.setup="score_data"} 
# point biserial 
iA <- itemAnalysis(response, hardFlag=.25, pBisFlag=.15)
iA$itemReport$pBis = round(iA$itemReport$pBis,2)
print(iA$itemReport)

# Plot items based on the pre-defined cut score (0.2)
item.discrimination <- data.frame(item = 1:20 ,
                                  discrimination =iA$itemReport$pBis)
cvpb = 0.2
plot(item.discrimination,
     type = "p",
     pch = 1,
     cex = 3,
     col = "purple",
     ylab = "Item-Total Correlation",
     xlab = "Item Number",
     ylim = c(0, 1),
     main = "Test Item Discriminations")

abline(h = cvpb, col = "red")

outlier <- data.matrix(subset(item.discrimination,
                              subset = (item.discrimination[, 2] <cvpb)))

text(outlier, paste("i", outlier[,1], sep = ""), col = "red", cex = .7)
```

### Item Difficulty and Discrimination `

```{r, icc2, echo=TRUE, exercise=TRUE, exercise.setup="score_data"} 
iA <- itemAnalysis(response, hardFlag=.25, pBisFlag=.15)
result = iA$itemReport 
plot(y=result$itemMean, x=result$pBis, 
    ylim=c(0,1) , xlim=c(0,1), 
     pch=18, 
     cex=2, 
     col="#69b3a2",
     xlab="Item Discrimination", ylab="Item Difficulty",
     main="A simple scatterplot"
     )

```

### Item Characteristic Curve ``cttICC``
- Given the information above, identify the two items with the highest and the lowest discrimination index.  
- Plot the item characteristic curve using \
`cttICC(data$score, response$#ItemNameHere, colTheme="spartans", cex=1.5)`
- How do the two plots differ?
```{r, icc, echo=TRUE, exercise=TRUE, exercise.setup="score_data"} 



```

## 4. Distractor Analysis 

Option/distractor analysis shows the percentage of students choosing each option in a dichotomous item.

```{r, distract, echo=TRUE, exercise=TRUE, exercise.setup="score_data"} 
# Distractor Analysis 
result <- distractor.analysis(CTTdata,CTTkey)
result
```
