---
title: "Week 5 - Reliability"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```

## Alternate-form Reliability 
```{r, af, echo=TRUE, exercise=TRUE}
data = read.table("./data/AF_data.dat",header=TRUE)
head(data)
assign("data", data, envir=globalenv()) 
```

### Score distribution of Form A and B 

```{r, af3, echo=TRUE, exercise=TRUE}
# compute the total score of Form A
FormA = 

# compute the total score of Form B 
FormB = 
# plot the score distribution of Form A using `hist()`
hist(FormA)

# plot the score distribution of Form B using `hist()`
hist(FormB)

#assign the variables as global variables 
assign("FormA", FormA, envir=globalenv()) 
assign("FormB", FormB, envir=globalenv())
```

### Calculate the Alternate-form Reliability 
- Between the form A and B `FormA` and `FormB`

```{r, af2, echo=FALSE, exercise=TRUE}
cor(FormA,FormB)
```

```{r, af2-solution}
cor(FormA,FormB)
```


## Test-Retest Reliability 

```{r, tr, echo=TRUE, exercise=TRUE}
data = read.table("./data/TR_data.dat",header=TRUE)
head(data)
assign("data", data, envir=globalenv()) 
```

### Score distribution of Test 1 and Test 2 
```{r, tr2, echo=FALSE, exercise=TRUE}
# compute the total score of Form A
Test1 = 

# compute the total score of Form B 
Test2 = 
# plot the score distribution of Form A using `hist()`
hist(Test1)

# plot the score distribution of Form B using `hist()`
hist(Test2)

#assign the variables as global variables 
assign("Test1", FormA, envir=globalenv()) 
assign("Test2", FormB, envir=globalenv())
```

### Calculate the Test-Retest Reliability 
- Between the Test 1 and Test 2 (`Test1` and `Test2`)

```{r, TR3, echo=FALSE, exercise=TRUE}
cor(Test1, Test2)
```

```{r, TR3-solution}
cor(Test1, Test2)
```


## Split-Half Reliability 
```{r, sh, echo=TRUE, exercise=TRUE}
data = read.table("./data/SH_data.dat",header=TRUE)
data = data[,2:61]
assign("data", data, envir=globalenv()) 
```

### Selecting Random Items For Halves 
```{r, sh2, exercise=TRUE, echo=TRUE}
# selecting random items for halves
halves = sample(1:60, 60, replace=F)
half1 = halves[1:30]
half2 = halves[31:60]

assign("half1", half1, envir=globalenv()) 
assign("half2", half2, envir=globalenv()) 
```

### Creating data set of half items 

```{r, sh3, exercise=TRUE, echo=TRUE}
# creating data set of half 1 items only and calculating summated scores
data1 = subset(data,select=half1)
X_1 = rowSums(data1)

# creating data set of half 2 items only and calculating summated scores
data2 = subset(data,select=half2)
X_2 = rowSums(data2)

assign("X_1", X_1, envir=globalenv()) 
assign("X_2", X_2, envir=globalenv()) 
```

### Correlation between the Two Scores 
```{r, sh4, exercise=TRUE, echo=FALSE}
# correlation between the two summated score variables
corr = cor(X_1,X_2)

# Spearman Brown's reliability between two split halves
rho = (2*corr)/(1+corr)
```

### Repeating the process 
```{r, sh5, exercise=TRUE, echo=FALSE, exercise.lines=35}
result =c()
for(y in 1:1000){
    # selecting random items for halves
    halves = sample(1:60, 60, replace=F)
    half1 = halves[1:30]
    half2 = halves[31:60]
    
    # creating data set of half 1 items only and calculating summated scores
    data1 = subset(data,select=half1)
    X_1 = rowSums(data1)
    
    # creating data set of half 2 items only and calculating summated scores
    data2 = subset(data,select=half2)
    X_2 = rowSums(data2)
    
    # correlation between the two summated score variables
    corr = cor(X_1,X_2)
    
    # Spearman Brown's reliability between two split halves
    rho = (2*corr)/(1+corr)
    
    # assigning a name to rho that has the repetition number in it
    name = paste("r",y,sep="")
    assign(name, value=rho)
    result = c(result, rho)
    # ending the loop
    }

# obtaining the average rho across the 1,000 repetitions
ymax <- max(result)
ymin <- min(result)

plot(result, type="l", col="#0000ff10", ylim=c(ymin, ymax))
abline(h=mean(result), col="blue")
```

## Cronbach's Alpha 

$$ \alpha = \frac{k}{k-1}\left(\frac{2\sum_{i\neq{j}}\sigma_{ij}}{\sigma_{c}^2}\right)\text{,} $$ 

```{r, ca, echo=TRUE, exercise=TRUE, exercise.lines= 23}
data = read.table("./data/Cronbach_data.dat",header=TRUE)
assign("data", data, envir=globalenv()) 

# calculating total summated scores
X = rowSums(data)

# calculating variance of total summated scores
varX = var(X)

# calculating variance of each item score
var_matrix = var(data)
vars = diag(var_matrix)

# calculating sum of item variances
sum_vars = sum(vars)

# obtaining number of items
n_items = length(vars)

# calculating alpha
alpha = (n_items/(n_items-1))*(1-(sum_vars/varX))

print(alpha)
```

#### Using the `CTT()` package fromt the last lection
```{r, ctt2, exercise=TRUE}

library(CTT)
reliability(data, itemal=TRUE)
```

#### Reliability and Confidence Interval 

If we know the mean and variance of the observed scores on a test, as well as the reported reliability, we can estimate a confidence interval around an estimated true score. Let’s use Alba’s predicted true score.


```{r ci, exercise=TRUE, echo=FALSE, exercise.lines=30}
rel <- 
SEM <- 
VarX <- 
Var_E  <- (1-rel)*VarX
SE_Msr <- sqrt(Var_E) 
LB     <- x - 1.96*SEM
UB     <- x + 1.96*SEM

``` 


## Spearman-Brown Prophecy

The Spearman-Brown prophecy formula is great for seeing how many more (or fewer) items we will need to achieve a desired reliability.

$$ \rho_{_{CC'new}} = \frac{k\rho_{_{CC'old}}}{1 + (k - 1)\rho_{_{CC'old}} } $$ 

For instance, how many more items do we need to achieve the reliability of 0.90 

$$ k = \frac{\rho_{_{CC'new}}(1 - \rho_{_{CC'old}})}{\rho_{_{CC'old}}(1 - \rho_{_{CC'new}}) }\\ $$ 

```{r, sp, exercise=TRUE}
alpha = 0.83
k = (0.90*(1-alpha))/(alpha*(1-0.90))

library(CTT)
relachieved_if_doubled <- spearman.brown(alpha, input = 1.8, n.or.r = "n")
relachieved_if_doubled
``` 
