---
title: "Lab Name"
author: "Author Name"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### Assignment 1: 
You’ve been provided operational data on a 10 item reading comprehension test taken by N= 831 5th grade students. We will call this `data`. All items are binary scored. This test is intended to be used across a wide range of student reading comprehension abilities for low-stakes formative purposes in the classroom. Complete and report the following analyses:

```{r}
data <- read.table('./data/timss_g4_2015_10_reading_items.csv')
data <- as.data.frame(data)
assign("data", data, envir=globalenv())
```

## Q1
*(2 points)* Provide a single table displaying frequency distributions of each item’s data. Compare the frequency distribution of item 7 to all other items.

```{r, q1, echo=FALSE}
###################### YOUR CODE HERE ##########################
data

# here you go

################################################################
```


## Q2

*(3 points)* Provide the CTT item difficulty estimate (p) for each item. 

- (1) Select an item that you consider to be relatively easy for the sample of 831 examinees. 
- (2) Would this be true for other samples from the same population? Why or why not?
- (3) Would this be true for samples from another population? Why or why not?

```{r, q2, echo=FALSE}
###################### YOUR CODE HERE ##########################
colSums(data)/831


###################### YOUR ANSWER HERE ##########################
# (1)
#
# (2)
#
# (3) 
#
###################################################################
```


## Q3

*(2 points)* Create a joint frequency distribution table for items 2 and 3. Calculate the inter-item correlation between item 2 and item 3, using the formula we defined in class. 

- Describe the bivariate relationship between the two items based on both the joint frequency distribution table and the correlation statistic.

```{r, q3, echo=FALSE}
###################### YOUR CODE HERE ##########################




###################### YOUR ANSWER HERE ##########################
# (1)
#
#
#
###################################################################
```

## Q4
*(2 points)* Create a joint frequency distribution table for items 3 and 5. Calculate the inter-item correlation between item 3 and item 5, using the formula we defined in class. 

- Compare the joint frequency distribution and correlation coefficient to those in item (c).

```{r, q4, echo=FALSE}
###################### YOUR CODE HERE ##########################




###################### YOUR ANSWER HERE ##########################
# (1)
#
# 
#
###################################################################
```

## Q5
*(2 points)* Calculate CTT item discrimination for each item. 

- Provide a summary interpretation of the discrimination estimates.)

```{r, q5, echo=FALSE}
###################### YOUR CODE HERE ##########################




###################### YOUR ANSWER HERE ##########################
# (1)
#
#
#
###################################################################
```


