---
title: "G-Theory"
author: "Jinnie shin | jinnie.shin@coe.ufl.edu"
date: "10/18/2021"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
#install.packages("hemp")
library(hemp)
library(learnr)
library(questionr)
```

## 1. Data 

G-theory expands our understanding of reliability to include multiple facets. The results of a G-study reveal how much variance in the scores is estimated to be explained by

- the object of measurement (usually persons),
- each of the facets,
  - including the interactions among facets, and
  - all interactions with the object of measurement except the highest-order interaction, and
- error (or residuals), which is confounded with the highest order interaction in the study.


```{r, data, exercise=TRUE, echo=TRUE}
library(hemp)
data(efData)
head(efData)

dat <- efData
```

Let’s turn the variables  into factor-type variables:
```{r, data_clean, exercise=TRUE, echo=TRUE, exercise.setup="data"}

dat$Participants <- factor(dat$Participants)
dat$Items <- factor(dat$Items)

str(dat)
```

## 2. One-facet study 

### G-Study 
Let's take a look at Task 1.  

For a one-facet study, say with raters who score persons, this would be the variance components of

- Persons ($\sigma_{p}^2$)
- Raters ($\sigma_{r}^2$)
- the person-by-rater interaction confounded with error ($\sigma_{pr,e}^2$)

```{r, r1, exercise=TRUE, echo=FALSE, exercise.setup="data_clean"}
fit   <- lm(Score ~ Participants + Items, data = dat)
anova1 <- anova(fit)
anova1 
``` 

In order to calculate the mean-squares...

```{r, r1-f, exercise=TRUE, echo=FALSE, exercise.setup='r1'}

MSp    <- (anova1["Participants",    "Mean Sq"])
MSr    <- (anova1["Items",     "Mean Sq"])
MSpr_e <- (anova1["Residuals", "Mean Sq"])
```

In order to calculate the variance components ... 

Because there are several raters, the  `MSp` right now includes the variance across the collections of these raters. Dividing by the number of raters, n_r, estimates this as the average variance component per rater, which is an estimate of the variance component in the population of raters (or universe of raters).

```{r, r1-f2, exercise=TRUE, echo=FALSE, exercise.setup='r1-f'}

n_p <- length(levels(dat$Participants))
n_r <- length(levels(dat$Items))

VCp <- (MSp - MSpr_e)/n_r #  the expected variance of the persons across a single condition of the other facet 
VCr <- (MSr - MSpr_e)/n_p
VCpr_e <- MSpr_e

```

### D-Study 
Now that we have the expected variance components, we can estimate the reliability (G-coefficient). The formula is
$$ G = \frac{\hat{\sigma}_{p}^2 }{\hat{\sigma}_{p}^2 + \hat{\sigma}_{REL}^2} $$
the error term $$ \hat{\sigma}_{REL}^2 $$ is all the sources of variation that have an effect on persons’ relative standing in the scores.

$$ \hat{\sigma}_{REL}^2 = \frac{\hat{\sigma}_{pr,e}^2}{n_r} $$ 

```{r, d1, exercise=TRUE, echo=FALSE, exercise.setup="r1-f2"}
Err_Rel <- VCpr_e/n_r
Err_Rel

G <- VCp/(VCp + Err_Rel)
round(G,3)

``` 

#### Finding an optimal number of conditions in a facet (items)
How will the G-coefficient change if we add or remove raters?
```{r, d2, exercise=TRUE, echo=FALSE, exercise.setup="r1-f2"}

# Let's select a range for the number of items 
Dn_i <- (1:15) # up to 15 items 

Err_Rel <- VCpr_e/Dn_i # relative error term 
G       <- VCp/(VCp + Err_Rel) # G coefficient 
Dstudy  <- data.frame(N_items = Dn_i, Err_Rel = round(Err_Rel,3), G_coef = round(G,3))
Dstudy

plot(Dstudy[,c("N_items", "G_coef")]) # plotting the results 
``` 

With this type of information, we can make decisions about the number of items to retain or add based on its effect on reliability 
