---
title: "Item Response Theory"
author: "Jinnie shin | jinnie.shin@coe.ufl.edu"
date: "10/18/2021"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
#install.packages("ltm")
library(ltm)
library(learnr)
library(questionr)
library(corrplot)
```

## 1. Data 

We will call the data (`Rasch_data.txt`) which includes 21 items answered by 200 students. 

```{r, data, exercise=TRUE, echo=TRUE}
data <- read.table("./data/Rasch_data.txt",header=FALSE,sep="")
head(data)
#Change the text file into matrix format that ltm can read
data <- as.matrix(data[1:200,1:20])
head(data)
```

## 2. Rasch Model 

### Running a rasch model 

```{r, rasch, exercise=TRUE, echo=FALSE, exercise.setup = "data"}

#Running rasch model 
#(the constraints are needed to hold all 
# discrimination parameters to 1, which 
# has to be true under the Rasch model)
RaschModel <- rasch(data, constraint=cbind(ncol(data)+1,1))
RaschModel
```

### Extracting estimated parameters 

```{r param, exercise=TRUE, echo=FALSE,  exercise.setup = "rasch"}
#Extracting estimated parameters
item.pars <- coef(RaschModel)
item.pars <- as.data.frame(item.pars)
item.pars
```

### Getting ability level 

```{r ability, exercise=TRUE, echo=FALSE,  exercise.setup = "param"}

#Getting ability estimates associated with each response vector
#(i.e., thetas are z values)
fscores <- factor.scores(RaschModel)

```

### Plotting 

```{r plot, exercise=TRUE, echo=FALSE , exercise.setup = "rasch"}

#Plotting ICCs for all items
plot(RaschModel, main="ICC from Rasch Model")

```

### Item Characteristic Curve 

$$p = exp((theta - b)) / (1 + exp((theta - b)) )$$

```{r, icc-value, exercise=TRUE, echo=FALSE, exercise.setup ="ability"}
latent = fscores$score.dat$z1

prob <- function(theta, b){
  exp(1.7*(theta - b)) / (1 + exp(1.7*(theta - b)) )
  
}
# the difficulty (b-param) of item 2 = 0.76944669
prob(theta =4, b=0.76944669)
  
```

#### Plotting ICC (item 2)

```{r, cal, exercise=TRUE, echo=FALSE, exercise.setup ="icc-value"}
p0 = prob(theta =-4, b=0.76944669)
p1 = prob(theta =-3, b=0.76944669)
p2 = prob(theta =-2, b=0.76944669)
p3 = prob(theta =-1, b=0.76944669)
p4 = prob(theta =0, b=0.76944669)
p5 = prob(theta =1, b=0.76944669)
p6 = prob(theta =2, b=0.76944669)
p7 = prob(theta =3, b=0.76944669)
p8 = prob(theta =4, b=0.76944669)

plot(c(p0, p1, p2, p3, p4, p5, p6, p7, p8), 
     x = c(-4, -3, -2, -1, 0, 1, 2, 3, 4), 
     xlab = "Ability", ylab="Probability")

```

#### Plotting for a single item 

```{r plot_one, exercise=TRUE, echo=FALSE,  exercise.setup = "rasch"}

#Plotting ICC for a single item
plot(RaschModel, items = c(2), main="ICC of Item 2")

```

### Item Information Function 
$$p = \frac{\exp^{1.7*(\theta - b)}}{1 + \exp^{1.7*(\theta - b)}}$$
$$ I(\theta) = p(1-p) $$

```{r, iif, exercise=TRUE, echo=FALSE, exercise.setup ="ability"}

info <- function(theta, b){
  p = exp(1.7*(theta - b)) / (1 + exp(1.7*(theta - b)) )
  p*(1-p)
}

b=0.76944669

info0 = info(theta = -3, b=b)
info1 = info(theta = -2, b=b)
info2 = info(theta = -1, b=b)
info3 = info(theta = 0, b=b)
info4 = info(theta = 1, b=b)
info5 = info(theta = 2, b=b)
info6 = info(theta = 3, b=b)
info7 = info(theta = 4, b=b)



plot(c(info0, info1, info2, info3, info4, info5, info6, info7),
     x = c(-4, -3, -2, -1, 0, 1, 2, 3),
     xlab = "Ability", ylab="Information")
```

### Plotting Item and Test Information Curve (TCC)

```{r icc, exercise=TRUE, echo=FALSE, exercise.setup = "rasch"}
#Plotting TIC
plot(RaschModel,  type=c("IIC"), main="Test Information Function")

#Plotting ICC
plot(RaschModel, items = c(2), type=c("IIC"), main="Item 2 Information Function")

```

```{r question, echo=FALSE}

question('What is the information for this item at θ = -4?', 
  answer("0.01", correct=TRUE), 
  answer("0.001"), 
  answer("0.0001"),
  answer("0.00001"))
```

### Plotting Test Information Curve (ICC)

```{r icc, exercise=TRUE, echo=FALSE, exercise.setup = "rasch"}
#Plotting TCC
plot(RaschModel,  type=c("IIC"), main="Item 2 Information Function")

```

### Item person map 
```{r, item-person, exercise=TRUE, echo=FALSE, exercise.setup = "ability"}
#Creating an item person map
plot(fscores, include.items = TRUE, main ="Item Person Map under Rasch model")

``` 

### Assessing item fit

null hypothesis is basically that observed data matches expected values
(more on item fit next week; really this should be checked prior to interpreting
all of the outputs above, such as difficulty estimates, ICCs, and TCC)

```{r fit, exercise=TRUE, echo=FALSE, exercise.setup = "rasch"}
item.fit(RaschModel,simulate.p.value = TRUE)
```


## 3. 2-PL and 3-PL 

### Data

Let's call the data (`2&3PL_data.txt`) which included a total of 3,000 students' responses on 20 dichotomously scored items. 

```{r, data2, exercise=TRUE, echo=FALSE}

temp <- read.table("./data/2&3PL_data.txt",header=FALSE,sep="")

#Change the text file into matrix format that ltm can read
data <- as.matrix(temp[1:3000,1:20])

```

### Running 2PL Model

```{r, 2pl, exercise=TRUE, echo=FALSE, exercise.setup = "data2"}

#Running 2-PL model
two.pl.model <- ltm(data ~ z1)

#Assessing item fit
#null hypothesis is that observed data matches expected values
item.fit(two.pl.model,simulate.p.value=TRUE)

#Extracting estimated parameters
item.pars.two.pl <- coef(two.pl.model)
item.pars.two.pl <- as.data.frame(item.pars.two.pl)
item.pars.two.pl

#Plotting ICCs 
plot(two.pl.model, main="ICC Curves under 2-PL Model")

#Creating an item person map
fscores <- factor.scores(two.pl.model)
plot(fscores, include.items = TRUE, main ="Item Person Map under 2PL model")
```

### Running 3PL Model

```{r, 3pl, exercise=TRUE, echo=FALSE, exercise.setup = "data2"}

#Running 3PL model 
three.pl.model <- tpm(data)

#Assessing item fit
#null hypothesis is that observed data matches expected values
item.fit(three.pl.model, simulate.p.value=TRUE)

#Extracting estimated parameters
item.pars.three.pl <- coef(three.pl.model)
item.pars.three.pl <- as.data.frame(item.pars.three.pl)
item.pars.three.pl

#Plotting ICCs
plot(three.pl.model, main="ICC Curves under 3-PL Model")

#Creating an item person map
fscores <- factor.scores(three.pl.model)
plot(fscores, include.items = TRUE, main ="Item Person Map under 3PL model")

```

